\begingroup
\vspace{3mm}
\abnormalparskip{0.5\baselineskip}
A generic flow system can be described as a system whose state depends on
flowing streams of energy, material, or information. Traditional examples of
such systems are the transport of properties such as pressure, temperature, and
matter in fluids, and the transport of charge in electrical currents. However,
valuable insight into various phenomena, including --- but not limited to ---
algal blooms in the ocean, and crowd patterns formed by humans, can be obtained
by regarding them as flow systems.

Lagrangian coherent structures (henceforth abbreviated to LCSs) can be
described as ``landscapes'' within a multidimensional phase space, which exert
a major influence upon the flow patterns in dynamical systems. In terms of
predicting future states of complex systems accurately, LCSs provide a simpler
means of analysis than the conventional approach of progressively increasing
the resoluton of the model(s) involved in numerical simulations. In this
context, a complex system is a system which exhibits sensitive dependence on
initial conditions. From their variational theory, hyperbolic LCSs are
identified as locally most repelling or attracting material surfaces. Somewhat
simplified, such surfaces can be considered as generalized trajectories,
created by the underlying transport phenomenon. When investigating transport
systems, hyperbolic LCSs are of particular interest, as they form the skeleton
of observable flow patterns.

Several important, naturally occuring transport phenomena --- such as the
spread of contaminants by oceanic surface currents --- can reasonably be
approximated as planar. This is one possible reason why previous
LCS research has mostly been conducted for two-dimensional systems; another may
be that the inclusion of a third dimension severely increases the underlying
mathematical complexity. To the extent that LCS analysis for three-dimensional
flows has hitherto been conducted, a common approach has been to merge a set
of two-dimensional projections, resulting in quasi-three-dimensional
surfaces. This project is centered around combining the underlying variational
principles of hyperbolic LCSs with an acknowledged technique for computing
invariant manifolds of vector fields, with a view to developing a robust
framework for computing hyperbolic LCSs in three-dimensional flow systems.

Our approach is described in detail, from its theoretical fundament, through
how we incorporated the underlying LCS theory into the computation of manifolds
and subsequently LCS surfaces, to some clever optimizations regarding
computational resource management. Using a few reference cases, we demonstrate
the efficacy of our method in reproducing simple three-dimensional manifolds
and LCSs. Moreover, the LCSs obtained for a simple, yet chaotic flow
system are shown to correspond well with those obtained for a time perturbed
version of the same system. This suggests that the computed LCS surfaces are
quite robust, and that the unavoidable inaccuracies of model data for
simulating transport phenomena need not be a hindrance for applying LCS
analysis to real-world systems.

It remains to be seen whether computing ``truly'' three-dimensional LCS surfaces
yields a sufficient increase in descriptive power, compared to the
aforementioned quasi-three-dimensional approach, to warrant the increase in
conceptual complexity and computational resource consumption. This balancing
act might be context sensitive.
\endgroup
