I would like to extend my gratitude to my supervisor, Assoc. Prof. Tor Nordam.
Our many discussions over the past academic year has provided me with great
insight, while his great knowledge and humour proved excellent tools in terms
of relaxing my early-semester nerves. Coupled with his expertise concering the
use of supercomputers for scientific purposes --- without the use of which,
none of the computations constituting this thesis would have been practically
feasible --- I could not be more satisfied.

My dear friend and fellow student Simon Nordgreen, who, much like me, thrived
under the guidance of Assoc. Prof. Nordam, definitely deserves a mention. Like
for our specialization projects which took place last semester, we have
collaborated closely for our master's thesis work. Our many thorough
discussions, and sessions of intensive work, have been invaluable to further
enhance my understanding of the underlying variational principles which form
the basis of our investigations. Looking back at what we have achieved over the
past year, I think it is safe to say that two heads are better than one.

%To all the friends and acquaintances I have had the pleasure of getting to know
%during my time in Trondheim, I extend my thanks. Our numerous collaborations
%and --- perhaps more importantly --- laughs have certainly helped brighten the
%many hours I have spent in study halls over the past few years.

Last, but most certainly not least, I gratefully thank my family for their
unconditional support and encouragement throughout.

\begin{minipage}[t]{\textwidth}
\begin{flushright}
A. M. T. L.
\end{flushright}
\end{minipage}
