\section[Macroscale stopping criteria for the expansion of computed manifolds]
{Macroscale stopping criteria for the expansion of \\\phantom{3.10} computed manifolds}
\label{sec:macroscale_stopping_criteria_for_the_expansion_of_computed_manifolds}

In principle, the process of developing manifold approximations by adding more
and more mesh points, organized in geodesic level sets, would continue as long
as the overall mesh quality was conserved (cf.\
\cref{sec:managing_mesh_accuracy}). The enforced (pseudo-)uniform expansion
radially outwards, inherent to the method of geodesic level sets, would then
yield a mesh providing a conservative estimate of the size of the
\emph{actual} manifold, as there is no particular reason to expect it to
expect the manifold to appear homogeneous, when viewed for the epicentre
of the computed level sets. With reasonable parameter choices, we became
able to generate large manifolds quite quickly. This lead us to enforce two
additional types of stopping criteria, based on what would happen if the
computed manifolds reached the edges of the computational domain, or folded
into themselves --- where the latter of the two will be described in the
imminent \cref{sub:continuous_self_intersection_checks}

The trajectories which are computed in order to identify new mesh points
(per the method described in
\cref{sec:revised_approach_to_computing_new_mesh_points}) frequently overstep
the domains within which the Cauchy-Green strain eigenvalues and -vectors
are defined in order to compute mesh points on or near the domain
boundaries. Thus, in order to resolve the behaviour of manifolds near the
edges of the domain of interest, the aforementioned strain characteristics
need to be computed in a region which \emph{contains} the domain of interest,
expanding beyond it in all directions. This is how we treated the case of
tidal flow in the Førde fjord (as outlined in
\cref{sec:flow_systems_defined_by_gridded_velocity_data}). For perfectly
periodic flow systems, such  as (either variant of) the ABC flow --- described
in \cref{sec:flow_systems_defined_by_analytical_velocity_fields} --- this
reduces to the trivial exercise of utilizing the inherent periodicity
(that is, provided that the computational domain is sufficiently large
to encompass at least one cycle along each direction).

\subsection{Continuous self-intersection checks}
\label{sub:continuous_self_intersection_checks}

Per LCS existence criterion \eqref{eq:lcs_condition_a}, there must be a uniquely
defined direction of strongest repellence everywhere along a repelling LCS.
Furthermore, our method of expanding manifolds by adding mesh points organized
in geodesic level sets (cf.\
\cref{sec:revised_approach_to_computing_new_mesh_points}) is based on
continuous expansions(pseudo-)radially outwards from the manifold epicentre.
Accordingly, the intersection of any manifold with itself was interpreted as
a nonphysical artefact of accumulated numerical error. For that reason, we
sought to terminate the expansion of a manifold when self-intersections
were detected.

Our method of detecting manifold self-intersections is based on comparing the
continuously computed interpolation triangles (as described in
\cref{sec:continuously_reconstructing_three_dimensional_manifold_surfaces_from_point_meshes})
--- in particular, we compared each triangle which was added together with the
most recently computed level set, to all of the triangles which had been
added with the preceding level sets. If at least one pair of triangles
intersected, the newest level set was flagged as self-intersecting. Our way
of determining whether or not two triangles intersect, is based on
the Möller-Trumbore ray-triangle intersection algorithm, with a detection
sensitivity parameter $\epsilon=10^{-8}$ \parencite{moller1997fast}. A visual
representation of our triangle-intersection algorithm is available in
\cref{fig:intersection_flowchart}.

\input{mainmatter/method/figures/intersection_algrthm_flowchart}

Some special cases of intersecting triangles warrant special treatment.
In particular, as our triangulation method (outlined in
\cref{sec:continuously_reconstructing_three_dimensional_manifold_surfaces_from_point_meshes})
is based on generating triangles which share sides with its neighbors, we
decided to allow triangles to intersect along the edges. Similarly, we allowed
for two triangles to be identical --- which might happen if a manifold
folds onto itself perfectly. Our algorithmic way of treating these cases
resulted in a theoretical false negative; namely, the case of two triangles
intersecting in exactly two points, laying along the edges of \emph{both}
triangles. In our experience, however, this never proved problematic ---
should one pair of triangles happen to intersect in this exact fashion
(which is a rarity in itself due to numerical round-off errors), another
pair of triangles would intersect in such a way that the most recently added
set as a whole would be flagged as self-intersecting. Two of the aforementioned
special cases --- that is, two triangles sharing an edge, and two triangles
intersecting in exactly two points laying along the edges of both triangles ---
are shown in \cref{fig:mollertrumbore_specialcases}.

\input{mainmatter/method/figures/moller_trumbore_specialcases}

Initial tests revealed that some self-intersections were quite innocuous, in
that, if the computed manifold was expanded by an additional level set,
the newly added triangulations need not necessarily intersect with any of the
preceding ones. This kind of insipid intersections could be a consequence of the
linear nature of our triangulation method, possibly compounded by round-off
error. Accordingly, we decided to stop the manifold expansion process
if \emph{several consecutive} geodesic level sets introduced intersection
triangulations. This was done by computing the sum of the interset distances
$\{\Delta_{i}\}$ for each consecutive level set $\{\mathcal{M}_{i}\}$ which
introduced new intersections. Whenever this pseudo-intersection length
exceeded $5\cdot\Delta_{\min}$ (cf.\ \cref{sub:maintaining_mesh_point_density}),
we terminated the manifold computation process; empirical trials suggested
that the intersection issue would then only worsen if more mesh points were
added.

