\section{Computed LCSs in the Førde fjord}
\label{sec:computed_lcss_in_the_forde_fjord}

Just like for the ABC flow, we used the $\mathcal{U}_{0}$ domain --- i.e.,
the grid points satisfying the LCS existence criteria
\eqref{eq:lcs_condition_a},~\eqref{eq:lcs_condition_b} and
\eqref{eq:lcs_condition_d} (the implementations of which are described in
\cref{sub:identifying_suitable_initial_conditions_for_developing_lcss}) ---
obtained for the transport system governed by the oceanic currents in the
Førde fjord as a first approximation to where repelling LCSs can reasonably
be expected to be found. Four different views of the $\mathcal{U}_{0}$ domain
are shown in \cref{fig:fjord_abd}. Compared to the corresponding domains
for the two variants of the ABC flow (as shown in
\cref{fig:steady_abd,fig:unsteady_abd}), the flow in the Førde fjord appears
a lot more chaotic, with fewer discernible macroscopic trends.
\Cref{fig:fjord_abd_y,fig:fjord_abd_x} do, however, indicate that the
points in the oceanic $\mathcal{U}_{0}$ domain are loosely organized
in horizontal layers.

\input{mainmatter/results/figures/fjord-abd.tex}

Using the filtering parameters provided in \cref{tab:initialconditionparams},
we identified initial conditions for the development of manifolds as a subset
of the $\mathcal{U}_{0}$ domain; yielding a total of \numprint{1631} points.
Then we computed manifolds and extracted LCSs using the parameters
in \cref{tab:initialconditionparams,tab:fjord_manifold_params} and the method
outlined in \cref{sec:preliminaries_for_computing_repelling_lcss_in_3d_flow%
    _by_means_of_geodesic_level_sets,%
    sec:revised_approach_to_computing_new_mesh_points,%
    sec:managing_mesh_accuracy,%
    sec:continuously_reconstructing_three_dimensional_manifold_surfaces_from_%
    point_meshes,%
    sec:macroscale_stopping_criteria_for_the_expansion_of_computed_manifolds,%
    sec:identifying_lcss_as_subsets_of_computed_manifolds}. This resulted in
a total of \numprint{110} LCS surfaces. As \cref{fig:fjord_lcss} indicates,
these largely appear to be organized in a sequence of horizontal layers.
Furthermore, all of the LCSs are sufficiently large to warrant treating them
as individual entities. Thus, in contrast to the LCSs obtained in (either
version of) the ABC flow (see \cref{sec:computed_lcss_in_the_abc_flow}), we
elected to assign each LCS a numerical value
\begin{equation}
    \label{eq:fjord_lcs_colorscaling}
    \mathcal{Q}_{i} = %
    \frac{\log{{\big({\mkern2mu}\overline{\lambda}_{3}\big)}_{i}}}%
    {\max\limits_{i}%
    \big\{\log{{\big({\mkern2mu}\overline{\lambda}_{3}\big)}_{i}}\big\}},
\end{equation}
where ${\big({\mkern2mu}\overline{\lambda}_{3}}\big)_{i}$, the repulsion
average of LCS surface $i$, is defined in \cref{eq:lcs_lm3_weight}. We then
used the (unit normalized) set of numbers $\big\{\mathcal{Q}_{i}\big\}$ to
select a color for the corresponding LCSs, drawn from a perceptually uniform
colormap.

\input{mainmatter/results/tables/fjord_params}

\input{mainmatter/results/figures/fjord-lcss.tex}

When comparing the computed LCSs (\cref{fig:fjord_lcss}) to the corresponding
$\mathcal{U}_{0}$ domain (\cref{fig:fjord_abd}), the organization of
the LCSs in horizontal layers seems reasonable. This is particularly apparent
from inspecting \cref{fig:fjord_abd_y,fig:fjord_lcss_y}, and
\cref{fig:fjord_abd_x,fig:fjord_lcss_x}. This indicates that the oceanic
flow undergoes the most stretching in the vertical direction. Moreover, the
repulsion appears to be largely uniform within any given depth layer.

Similarly to our treatment of an LCS surface in the steady ABC flow in
\cref{sub:verifying_that_the_computed_lcss_are_in_fact_repelling}, we placed
two blobs of particles on either side of the most strongly repulsive LCS
present in the fjord subdomain --- perhaps most easily spotted in the middle
of \cref{fig:fjord_lcss_y,fig:fjord_lcss_x} --- and allowed the oceanic
currents to transport the particles as well as the computed LCS for the
entirety of our \numprint{12} hour time interval of interest. The initial and
final states are shown in \cref{fig:blobtest-fjord}. Like for the ABC flow
case, the triangulated structure of the LCS breaks down, yet the two blobs of
particles diverge from each other in the transition from
\cref{fig:blobtest-fjord-pre} to \cref{fig:blobtest-fjord-post}. The particles
belonging to either blob remain close together, and none of them appear to
move across the LCS mesh point. This suggests that, like the ones computed
for the ABC flow, the LCSs obtained for flow in the Førde fjord act as
repelling material surfaces, and thereby also as transport barriers.

\input{mainmatter/results/figures/blobtest-fjord}


