Our method for computing repelling LCSs in three-dimensional flow systems
seems to yield reasonable results. Analytically constructed test cases indicate
that our method of generating three-dimensional surfaces, in addition to
extracting repelling LCSs as subsets of computed manifolds which satisfy all
the necessary and sufficient existence criteria, work as intended. Moreover,
the robustness of the LCSs obtained in the two variations of the
Arnold-Beltrami-Childress flow considered here, suggests that the computed LCSs
are not particularly sensitive to imperfect model data --- which is a
characteristic property of \emph{Lagrangian} transport barriers. Accordingly,
our computed LCSs for flow in the Førde fjord likely form reasonable
approximations of the actual LCSs contained within.

There is certainly room for further research with regards to the numerical
implementation of one of the LCS existence conditions derived from their
variational theory; in particular, that pertaining to the identification of
locally most repelling material surfaces (the other conditions are quite
unambiguous, in comparison). To our knowledge, a general, robust numerical
routine is yet to be described in the literature. For instance, in a recent
study conducted by \textcite{oettinger2016autonomous}, who set out to compute
quasi-three-dimensional hyperbolic LCSs, the authors seemingly did not attempt
to implement the aforementioned criteria numerically, appearing to be satisfied
with reasonable suggestions as to where such LCSs \emph{might} be located. Our
approach, based on identifying which of the individual points constituting each
manifold was locally repelling, involved one parameter which was determined by
use of the  initial grid spacing, and another which filtered away the smallest
LCS  surfaces --- as these were assumed not to impact the overall circulation
significantly (as originally suggested by \textcite{farazmand2012computing}).
A suggested alternative approach, which was not investigated as part of this
project, would be to utilize a sort of numerical clustering algorithm to
extract LCSs as the most repelling material surfaces in small neighborhoods,
where each of the considered sets of surfaces ideally would be of similar size
and orientation.

Seeing as our method of computing fully three-dimensional LCSs is significantly
more complex and consumes more computational resources than well established
routines for generating their two-dimensional counterparts (see e.g.\
\textcite{onu2015lcstool}), strong arguments regarding the additional insight
from three-dimensional anlysis are needed in order to substantiate a change in
practices. Moreover, computing fully three-dimensional LCSs might not be
necessary, depending on the type of transport phenomenon under consideration.
For instance, regarding the spread of debris and contaminations such as garbage
patches or oil spill remnants across the ocean surface, the underlying
transport system can reasonably be considered two-dimensional. Meanwhile, for
circulation in rivers or fjords, where the depth is of the same scale as the
width and vertical transport is often significant, fully three-dimensional
analysis might be more suitable. Simply put, for systems in which all three
dimensions are of similar relevance --- i.e.,\ systems which cannot reasonably
be regarded as two-dimensional --- computing fully three-dimensional LCS
surfaces might be prudent; in contrast to the quasi-three-dimensional approach
of \textcite{blazevski2014hyperbolic} which loses out on the minute details
of the extra dimension, and does not yield coherent surfaces favorable for
further inquiries.

\subsubsection{Suggestions for further work}
\label{ssub:suggestions_for_further_work}

Experimental methods for the verification of Lagrangian flow analysis have
gained traction lately. Investigations are currently being conducted as to the
usefulness of LCSs as predictors for diffusive flow patterns, for which
analyzing tracer trajectories of the overall circulation might be insufficient
\parencite{haller2018material}. Moreover, recent pilot studies indicate that
Lagrangian analysis yields valid predictions for transport by oceanic currents
\parencite{filippi2018detection} --- including the fully three-dimensional
circulation patterns arising in off-shore currents
\parencite{peacock2018targeted}. Accordingly, empirical studies pertaining to
the validity of the LCSs computed by use of our variation of the geodesic level
set method lies within the realm of feasibility.

In addition to the aforementioned issue of implementing the LCS existence
criterion which identifies LCSs as surfaces which are locally most repelling,
further research aiming to enhance our approach for computing repelling LCSs
could result in increased transparency and efficiency. Specifically,
abandoning the strict ordering of mesh points in level sets forming topological
circles could allow for an increase in the resolution of LCS behaviour near
domain boundaries, and potentially yield more accurate surface approximations
overall. Then again, other options for computing invariant manifolds of
three-dimensional vector fields exist. Aside from the method of geodesic level
sets, \textcite{krauskopf2005survey} present four others; investigating each of
them in the context of computing hyperbolic LCSs could provide valuable
insights. Moreover, a method less demanding in terms of resource
consumption than the present approach would render the computation of
three-dimensional LCSs a feasible task without the need of supercomputers.
Courtesy of facilitating three-dimensional Lagrangian analysis for a wider
audience, this could, in turn, accelerate the further development of LCS
computing tools as a whole.

Lastly, to our knowledge, little research has been conducted regarding the
choice of interpolation and integration methods in order to simulate transport
phenomena in real-world systems. High-order adaptive step size integration
methods are generally used in conjunction with interpolated velocity fields,
seemingly with little awareness as to whether or not the use of higher-order
methods are warranted. For instance, \textcite{vansebille2018lagrangian}
provide a rigorous discussion of the currently available tools for computer
simulated tracer advection, yet do not treat integration methods in detail.
The significance of the choice of interpolation scheme is generally recognized,
though left largely unexplored. Whereas \textcite{lekien2005tricubic} developed
a (locally) tricubic interpolation method intended for use in simulating
three-dimensional flow (and computing three-dimensional LCSs) --- motivated by
the observation that linear interpolation is largely insufficient in terms of
yielding smooth velocity fields from model data --- the authors provide no
arguments for preferring cubic methods over even higher order alternatives. A
third example is the article by \textcite{gough2017persistent}, in which the
authors report having used a high-order adaptive step size Runge-Kutta solver
together with a cubic interpolation routine --- much like what was done for
this project --- in order to investigate oceanic transport patterns in the
northwestern Gulf of Mexico; yet fail to motivate their choices of integration
and interpolation methods. Based on the above, investigations with regards to
the interaction between integration and interpolation schemes in the context
of computing LCSs are appealing, due to their innate relation to application.

